---
title: "Covid19 Situation in Outaouais"
output:
  html_document:
    theme: darkly
    toc: true
    toc_depth: 1
    includes:
      before_body: html/translate.html
params: 
  optimize: false
---

```{r setup, include=FALSE}
source("R/setup.R")
load("data/covid_local.RData")
load("data/covid_local_daily.RData")
load("data/municipalities.RData")
load("data/hospitalization.RData")
load("data/inspq.RData")
load("data/rls.RData")
rls <- rls %>% arrange(key, time)
load("data/vaccination.RData")
vaccination <- vaccination %>% 
  mutate(date = lubridate::as_date(vaccination$time)) %>% 
  group_by(key, date) %>% filter(time == max(time)) 
pop_outaouais_2020 <- 401388
source("R/functions_figures.R")
```

```{r, child = 'md/index_intro.md'}
```

# Outaouais region totals {.tabset .tabset-fade .tabset-pills}

## New cases
```{r cases_avg}
## DailyFig(c("Average increase per day (Outaouais)"), tab = "cases", per_pop = TRUE)
RegionFig(df = inspq[inspq$key == "Average increase per day", ], 
          bars = inspq[inspq$key == "New cases", ], 
          pop = "per100k",
          caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/covid19-hist.csv")
```

## Active cases
```{r cases_active}
## LayeredFig(c("Active cases"))
RegionFig(df = inspq[inspq$key %in% c("Active cases"), ], 
          pop = "per100k", 
          caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/covid19-hist.csv")
```

## Cumulative cases
```{r cases_cumulative}
## LayeredFig(c("Healed/resolved cases", "Total cases (Outaouais)")) # "Total cases (RLS)"
RegionFig(df = inspq[inspq$key %in% c("Total cases", "Total recovered cases"), ], 
          caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/covid19-hist.csv")
```

## Deaths
```{r cases_deaths}
## LayeredFig(c("Total deaths"))
RegionFig(df = inspq[inspq$key %in% c("Total deaths", "Total deaths (CHSLD)", "Total deaths (RPA)", "Total deaths (other)"), ], 
          caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/covid19-hist.csv")
```

## Hospitalizations
```{r hospitalization}
RegionFig(df = hospitalization[hospitalization$key %in% c("Active hospitalizations, non-ICU", "Active hospitalizations, ICU"), ],
          second = inspq[inspq$key %in% c("New hospitalizations, non-ICU", "New hospitalizations, ICU"), ], 
          caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/covid19-hist.csv &\n www.donneesquebec.ca/recherche/dataset/covid-19-portrait-quotidien-des-hospitalisations")
```

## Total hospitalizations
```{r hospitalizations}
RegionFig(df = inspq[inspq$key %in% c("Total hospitalizations, non-ICU", "Total hospitalizations, ICU", "Total hospitalizations"), ], 
          caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/covid19-hist.csv")
```

## Vaccination
```{r vaccination}
## RegionFig(df = vaccination,
##         caption = "Data source: www.quebec.ca/en/health/health-issues/a-z/2019-coronavirus/situation-coronavirus-in-quebec/covid-19-vaccination-data/")
RegionFig(df = inspq[inspq$key %in% c("Vaccine doses administered", "Total vaccine doses administered"), ], 
          pop = "percent", 
          dlabels = "%d %b %Y",
          caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/covid19-hist.csv")
```

# Outaouais & Ottawa (opencovid.ca) {.tabset .tabset-fade .tabset-pills}

```{r, child = 'md/note_ottawa.md'}
```

## Average increase
```{r opencovid_avg}
DailyFig(c("Average increase per day (Outaouais)", "Average increase per day (Ottawa)"), 
         tab = "opencovid.ca", 
         rug = FALSE, 
         caption = "Data source: opencovid.ca/api/")
```

## Cumulative cases
```{r opencovid_total}
LocalFig(c("Total cases (Outaouais)", "Total cases (Ottawa)"), 
         tab = c("opencovid.ca"), 
         rug = FALSE, 
         caption = "Data source: opencovid.ca/api/")
```

## Deaths
```{r deaths_opencovid}
LocalFig(c("Total deaths (Outaouais)", "Total deaths (Ottawa)"), 
         tab = c("opencovid.ca"), 
         rug = FALSE, 
         caption = "Data source: opencovid.ca/api/")
```

# Testing in Outaouais {.tabset .tabset-fade .tabset-pills}

## Daily Tests
```{r testing}
## LayeredFig(c("Average screening tests per day"))
RegionFig(df = inspq[inspq$key %in% c("Average testing per day"), ],
          bars = inspq[inspq$key %in% c("Persons tested"), ],
          caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/covid19-hist.csv")
```

## Positive tests
```{r positive}
RegionFig(df = inspq[inspq$key %in% c("Average positive tests per day"), ], 
          bars = inspq[inspq$key %in% c("Persons tested positive"), ],
          caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/covid19-hist.csv")
```

## Test positivity
```{r positivity}
RegionFig(df = inspq[inspq$key %in% c("Average test positivity (%)"), ], 
          bars = inspq[inspq$key %in% c("Test positivity (%)"), ], 
          caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/covid19-hist.csv")
```

# Average increase & cumulative cases by RLS {.tabset .tabset-fade .tabset-pills}

## Gatineau & Outaouais
```{r rls_avg_gatineau}
## DailyFig(c("RLS de Gatineau"))
RLSFig(df = rls[rls$key %in% c("Average increase per day (Outaouais)", "Active cases (Outaouais)", "RLS de Grande-Rivière - Hull - Gatineau") & rls$table == "average", ], 
       caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/tableau-rls-new.csv")
```

## Collines
```{r rls_avg_collines}
## DailyFig(c("RLS des Collines-de-l'Outaouais"))
RLSFig(df = rls[rls$key == "RLS des Collines-de-l'Outaouais" & rls$table == "average", ], 
       caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/tableau-rls-new.csv")
```

## Pontiac
```{r rls_avg_pontiac}
## DailyFig(c("RLS du Pontiac"))
RLSFig(df = rls[rls$key == "RLS du Pontiac" & rls$table == "average", ], 
       caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/tableau-rls-new.csv")
```

## Papineau
```{r rls_avg_papineau}
## DailyFig(c("RLS de Papineau"))
RLSFig(df = rls[rls$key == "RLS de la Vallée-de-la-Lièvre et de la Petite Nation" & rls$table == "average", ], 
       caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/tableau-rls-new.csv")
```

## Vallée
```{r rls_avg_vallee}
## DailyFig(c("RLS de la Vallée-de-la-Gatineau"))
RLSFig(df = rls[rls$key == "RLS de la Vallée-de-la-Gatineau" & rls$table == "average", ], 
       caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/tableau-rls-new.csv")
```

## Cumulative, Gatineau
```{r rls_cumulative_gatineau}
RLSFig(df = rls[rls$key %in% c("Total cases (Outaouais)", "RLS de Grande-Rivière - Hull - Gatineau") & rls$table == "cases", ], 
       caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/tableau-rls-new.csv")
```

## Cumulative, other
```{r rls_cumulative_other}
RLSFig(df = rls[rls$key %in% c("RLS des Collines-de-l'Outaouais", "RLS du Pontiac", "RLS de la Vallée-de-la-Lièvre et de la Petite Nation", "RLS de la Vallée-de-la-Gatineau") & rls$table == "cases", ], 
       caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/tableau-rls-new.csv")
```

# Active cases by RLS {.tabset .tabset-fade .tabset-pills}

## Gatineau & Outaouais
```{r rls_active_gatineau}
## LocalFig(c("RLS de Gatineau"), tab = c("rls_active"), legend_cols = 3)
RLSFig(df = rls[rls$key %in% c("Active cases (Outaouais)", "RLS de Grande-Rivière - Hull - Gatineau") & rls$table == "active", ], 
       caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/tableau-rls-new.csv")
```

## Collines
```{r rls_active_collines}
## LocalFig(c("RLS des Collines-de-l'Outaouais"), tab = c("rls_active"), legend_cols = 3)
RLSFig(df = rls[rls$key == "RLS des Collines-de-l'Outaouais" & rls$table == "active", ], 
       caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/tableau-rls-new.csv")
```

## Pontiac
```{r rls_pontiac}
## LocalFig(c("RLS du Pontiac"), tab = c("rls_active"), legend_cols = 3)
RLSFig(df = rls[rls$key == "RLS du Pontiac" & rls$table == "active", ], 
       caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/tableau-rls-new.csv")
```

## Papineau
```{r rls_active_papineau}
## LocalFig(c("RLS de Papineau"), tab = c("rls_active"), legend_cols = 3)
RLSFig(df = rls[rls$key == "RLS de la Vallée-de-la-Lièvre et de la Petite Nation" & rls$table == "active", ], 
       caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/tableau-rls-new.csv")
```

## Vallée
```{r rls_active_vallee}
## LocalFig(c("RLS de la Vallée-de-la-Gatineau"), tab = c("rls_active"), legend_cols = 3)
RLSFig(df = rls[rls$key == "RLS de la Vallée-de-la-Gatineau" & rls$table == "active", ], 
       caption = "Data source: www.inspq.qc.ca/sites/default/files/covid/donnees/tableau-rls-new.csv")
```

# Cumulative cases by municipalities {.tabset .tabset-fade .tabset-pills}

## Total & Gatineau
```{r areas_gatineau}
LocalFig(c("Gatineau", "Total cases (municipalities)"))
```

## Collines
```{r areas_collines}
LocalFig(municipalities$municipality[municipalities$mrc == "Les Collines-de-l'Outaouais"])
```

## Papineau
```{r areas_papineau}
exclude_papineau <- c("Thurso", "Papineauville", "Saint-André-Avellin", "Chénéville","Ripon", "Montebello")
LocalFig(keys = exclude_papineau)
```

## Papineau (<20)
```{r areas_papineau_cont}
LocalFig(municipalities$municipality[municipalities$mrc == "Papineau" & 
                                     !municipalities$municipality %in% exclude_papineau])
```

## Vallée
```{r areas_vallee}
exclude_vallee <- c("Messines", "Déléage", "Bouchette")
LocalFig(c(municipalities$municipality[municipalities$mrc == "La Vallée-de-la-Gatineau" & 
                                       municipalities$designation == "Ville"], exclude_vallee))
```

## Vallée (<20)
```{r areas_vallee_cont}
LocalFig(municipalities$municipality[municipalities$mrc == "La Vallée-de-la-Gatineau" & 
                                     municipalities$designation != "Ville" & 
                                     !municipalities$municipality %in% exclude_vallee])
```

## Pontiac
```{r areas_pontiac}
LocalFig(municipalities$municipality[municipalities$mrc == "Pontiac"])
```

# Information on this site {.tabset .tabset-pills}

```{r, child = 'md/index_caveats.md'}
```

```{r, child = 'md/index_notes.md'}
```

```{r, child = 'md/updates.md'}
```

Last data revision: `r str_replace(read_file("data/data_update_time.txt"), "\n", "")`

Last page update: `r now()`

```{r, child = 'md/table_glossary.Rmd', eval = FALSE}
```

```{r, child = 'md/table_sources.Rmd'}
```

```{r, child = 'md/disclaimer.md'}
```

```{r, child = 'md/footer.md'}
```
