---
title: "Covid19 Situation in Outaouais"
author: "Oskar Timo Thoms"
date: "11/03/2021"
output: 
  github_document: 
    toc: true
    fig_width: 10
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
knitr::opts_chunk$set(echo = TRUE)
```
```{r data, include=FALSE}
load("data/covid_local.RData")
# tapply(covid$key, covid$table, unique)
keys <- list(
    cases = list(
        c("Cumulative cases", "Healed/resolved cases"),
        c("Active cases", "Deaths", "Average screening test per day") # "Daily increase"
    ),
    rls = list(
        c("Active cases", "Total cases (active)", "Deaths"),
        c("Total cases", "Healed/resolved cases"),
        c("To be determined", "To be determined (active)"),
        c("RLS de Gatineau", "RLS de la Vallée-de-la-Gatineau", "RLS de Papineau", "RLS des Collines-de-l'Outaouais", "RLS du Pontiac"),
        paste(c("RLS de Gatineau", "RLS de la Vallée-de-la-Gatineau", "RLS de Papineau", "RLS des Collines-de-l'Outaouais", "RLS du Pontiac"), "(active)")
    ), 
    areas = list(
        c("Active cases", "To be determined"), 
        c("Total cases", "Healed/resolved cases"), 
        unique(covid$key[covid$table == "areas"])[!unique(covid$key[covid$table == "areas"]) %in% c("Active cases", "To be determined", "Total cases", "Healed/resolved cases")]
    )
)
```

It is difficult to find covid19 trend data at the local level for the Outaouais. The [Quebec government](https://www.quebec.ca/en/health/health-issues/a-z/2019-coronavirus/situation-coronavirus-in-quebec/) shows summaries but only a few days of trend data for regions. [Quebec Public Health](https://www.inspq.qc.ca/covid-19/donnees) shows longer time series but only cumulative snapshots by region. The data aggregated by the [COVID-19 Canada Open Data Working Group](https://opencovid.ca/) can be used to show trends at the health region level. None of these sources provide trend data below the regional level. [CISSS Outaouais](https://cisss-outaouais.gouv.qc.ca/language/en/covid19-en/) provides frequent (sometimes daily) snapshots by Réseaux locaux de services (RLS) and by municipality, but no trends. 

The present project provides local trend data based on these snapshots. The HTML code for the [CISSS Outaouais site](https://cisss-outaouais.gouv.qc.ca/language/en/covid19-en/) is downloaded daily, the tables are scraped and processed into a tidy dataset, to produce the figures on this site. The R code for downloading and processing is made available, but not the software for the daily automation. (This is currently done with Macscheduler and AppleScript.) Note several important caveats when consulting the trend data. 
 
## Caveats

1. There is no guarantee of data accuracy. I am aggregating what CISSS Outaouais has been reporting over time. While I do try to correct obvious data input error, I have (so far) not implemented an automatic error detection process.

2. The date and time reported in the dataset refer to when the CISSS Outaouais website was accessed, not when cases occurred. 

3. The trend data are not complete for several reasons. First, I started regularly downloading the CISSS Outaouais website in fall 2020, and used the [Wayback Machine](https://archive.org/web/) to get earlier snapshots. Unfortunately, due to a nasty syncing glitch, I lost many of the earlier HTML files. (RStudio projects and Dropbox do not play well together; I learnt this the hard way, twice.) Second, the CISSS Outaouais site is not updated every day. Third, running the code from my personal computer means that the automatic download does not happen every day, either because the computer is turned off, or because my software for scheduling scripts has failed a few times. Since the snapshots include cumulative counts, this is not necessarily a problem, but it means that changes in case counts cannot always be precisely dated. 

4. Not all the data available on the CISSS Outaouais site are included in the dataset.

5. Finally, at the municipal and RLS levels, CISSS Outaouais does not report numbers of less than 6 precisely, likely for good privacy reasons. When "5 or less" are reported, I record this as 5 cases. Consumers of the trend data must be aware that counts of 5 do not precisely reflect the actual situation at those levels, but refer to 1-5 cases.

Figures updated: `r now()`.

## Outaouais Region

```{r figures, echo=FALSE, results='hide'}
lapply(keys$cases, function(keys) {
    list(ggplot(data = covid[covid$key %in% keys, ]) + 
             geom_line(mapping = aes(x = time, y = value, group = key, color = key)) + 
             theme_bw() + labs(x = "", y = "") + 
             theme(legend.position = "bottom") + 
             ggtitle(paste(keys, collapse = " & ")),
         ggplot(data = covid[covid$key %in% keys & covid$time > now() - days(42), ]) + 
             geom_line(mapping = aes(x = time, y = value, group = key, color = key)) + 
             theme_bw() + labs(x = "", y = "") + 
             theme(legend.position = "bottom") + 
             ggtitle(paste(paste(keys, collapse = " & "), "(last 6 weeks)"))
    )
})
```

## By RLS

TBA

## By Municipalities

TBA
